### 2017-03-15 notes.tex
Analysis using the new data.
A.0.0. The initial analysis contains plots for measures of the entire
data set. This was done to investigate if anything stands out.
The enviroments still look cooperative.

A.1. Is the analysis at the tournament level of different types.
Now we are ploting the transitions to C.

Everything is still in primary stage.

Last notebook contains the note to myself for
getting back.

The notes.tex contains some notes for the variables.

Overall look, on how cooperation affects performance. In the experiment that was help,
the meta data of given tournaments have been collected. The metadata include:

- cooperation rating $(C_{r})$
- median score
- rank $(r_{i})$
- wins
- cc_rate
- cd_rate
- dc_rate
- dd_rate
- noise
- probend
- repetitions
- seed
- size
- turns.

In this section we will rely on the rank to monitor the perfomance of the individuals within the tournaments and hopefully get some insights on our tournaments. More specificaly we will use the normalized rank $\bar{r_{i}} = \frac{r_{i}}{N_{i}}$, because there are different tournament sizes.

We will make use of some other mathematical representations as well, such as, the
index of the tournament $i$ and the number of players $N_{i}$.