### 2019-10-24 paper/top_perfomances.tex
draft people hanging it to Vince

\begin{tabular}{lggllggllr}
\toprule
& \multicolumn{2}{g}{Standard} & \multicolumn{2}{c}{Noisy} & \multicolumn{2}{g}{Probabilistic ending} &  \multicolumn{2}{c}{Noisy probabilistic ending} \\
\midrule
& Name & $\bar{r}$ &                 Name & $\bar{r}$ &               Name & $\bar{r}$ &                 Name & $\bar{r}$ \\
\midrule
1 &           Evolved HMM 5 &   0.00667 &               Grumpy &    0.1402 &          Fortress4 &   0.01266 &           Alternator &    0.3037 \\
2 &          Evolved FSM 16 &   0.00995 &                  $e$ &   0.19388 &           Defector &   0.01429 &               $\phi$ &   0.30978 \\
3 &    EvolvedLookerUp2 2 2 &   0.01064 &       Tit For 2 Tats &    0.2069 &  Better and Better &   0.01587 &                  $e$ &    0.3125 \\
4 & Evolved FSM 16 Noise 05 &   0.01667 &         Cycle Hunter &   0.21538 &    Tricky Defector &   0.01875 &                $\pi$ &   0.31686 \\
5 &       PSO Gambler 2 2 2 &   0.02143 &       Risky QLearner &   0.22222 &          Fortress3 &   0.02174 &    Limited Retaliate &   0.35263 \\
6 &             Evolved ANN &   0.02878 &          Retaliate 3 &   0.22887 &     Gradual Killer &   0.02532 &     Anti Tit For Tat &   0.35431 \\
7 &           Evolved ANN 5 &    0.0339 &        Cycler CCCCCD &   0.23507 &         Aggravater &   0.02778 &          Retaliate 3 &   0.35563 \\
8 &       PSO Gambler 1 1 1 &   0.03704 &          Retaliate 2 &   0.23913 &             Raider &   0.03077 &  Limited Retaliate 3 &   0.35563 \\
9 &           Evolved FSM 4 &   0.04891 &      Defector Hunter &   0.24038 &         Cycler DDC &   0.04545 &            Retaliate &   0.35714 \\
10 &        PSO Gambler Mem1 &   0.05036 &            Retaliate &   0.24177 &        Hard Prober &   0.05128 &          Retaliate 2 &   0.35767 \\
11 &                Winner12 &   0.06011 &  Hard Tit For 2 Tats &      0.25 &         SolutionB1 &   0.06024 &  Limited Retaliate 2 &   0.36134 \\
12 &            Fool Me Once &    0.0614 &             ShortMem &   0.25286 &      Meta Minority &   0.06077 &             Hopeless &   0.36842 \\
13 &                     DBS &   0.07143 &  Limited Retaliate 3 &   0.25316 &              Bully &   0.06081 &    Arrogant QLearner &   0.40651 \\
14 &           DoubleCrosser &     0.072 &    Limited Retaliate &   0.25706 &    Fool Me Forever &    0.0708 &    Cautious QLearner &   0.40909 \\
15 &             BackStabber &   0.07519 &                $\pi$ &   0.25801 &             EasyGo &   0.07101 &      Fool Me Forever &   0.41764 \\
\bottomrule
    \end{tabular}