### 2020-01-28 paper/perfomances_table.tex
Prepares paper (#13)
* re-run analysis with new data

* clean repo

* move notebooks

* rename files, run black and tweak code

* tests to pytest

* test and implement prepare, process src files

* create clystering analysis script

* update enviroment yml

* add feautures to analysis

* add lines to run on siren

* tweaks to outputs

* add new meausures to analysis

* add probend, noise and turns to cluster analysis

* fix failure

* remove turns as default cluster feauture

* week commit

* write first sections of papers

* fix errors and move tests

* create test file

* tweats to clusterin

\begin{tabular}{lllllllll}
\toprule
{} &                     Name & \$\textbackslash bar\{r\}\$ &                   Name & \$\textbackslash bar\{r\}\$ &               Name & \$\textbackslash bar\{r\}\$ &                 Name & \$\textbackslash bar\{r\}\$ \\
\midrule
0  &            Evolved HMM 5 &   0.00667 &                 Grumpy &    0.1402 &          Fortress4 &   0.01266 &           Alternator &    0.3037 \\
1  &           Evolved FSM 16 &   0.00995 &                    \$e\$ &   0.19388 &           Defector &   0.01429 &               \$\textbackslash phi\$ &   0.30978 \\
2  &     EvolvedLookerUp2\_2\_2 &   0.01064 &         Tit For 2 Tats &   0.20617 &  Better and Better &   0.01587 &                  \$e\$ &    0.3125 \\
3  &  Evolved FSM 16 Noise 05 &   0.01667 &  Slow Tit For Two Tats &   0.20962 &    Tricky Defector &   0.01875 &                \$\textbackslash pi\$ &   0.31686 \\
4  &        PSO Gambler 2\_2\_2 &   0.02143 &           Cycle Hunter &   0.21538 &          Fortress3 &   0.02174 &    Limited Retaliate &   0.35263 \\
5  &              Evolved ANN &   0.02878 &         Risky QLearner &   0.22222 &     Gradual Killer &   0.02532 &     Anti Tit For Tat &   0.35431 \\
6  &            Evolved ANN 5 &    0.0339 &            Retaliate 3 &   0.22887 &         Aggravater &   0.02778 &          Retaliate 3 &   0.35563 \\
7  &        PSO Gambler 1\_1\_1 &   0.03704 &          Cycler CCCCCD &   0.23507 &             Raider &   0.03077 &  Limited Retaliate 3 &   0.35563 \\
8  &            Evolved FSM 4 &   0.04891 &            Retaliate 2 &   0.23913 &         Cycler DDC &   0.04545 &            Retaliate &   0.35714 \\
9  &         PSO Gambler Mem1 &   0.05036 &        Defector Hunter &   0.24038 &        Hard Prober &   0.05128 &          Retaliate 2 &   0.35767 \\
10 &                 Winner12 &   0.06011 &              Retaliate &   0.24177 &         SolutionB1 &   0.06024 &  Limited Retaliate 2 &   0.36134 \\
11 &             Fool Me Once &    0.0614 &    Hard Tit For 2 Tats &      0.25 &      Meta Minority &   0.06077 &             Hopeless &   0.36842 \\
12 &                      DBS &   0.07143 &               ShortMem &   0.25286 &              Bully &   0.06081 &    Arrogant QLearner &   0.40651 \\
13 &            DoubleCrosser &     0.072 &    Limited Retaliate 3 &   0.25316 &    Fool Me Forever &    0.0708 &    Cautious QLearner &   0.40909 \\
14 &              BackStabber &   0.07519 &      Limited Retaliate &   0.25706 &             EasyGo &   0.07101 &      Fool Me Forever &   0.41764 \\
\bottomrule
\end{tabular}