### 2020-03-27 paper/parameters_section.tex
finalise introduction

\section{Parameters Summary}\label{app:parameters}

All the parameters used in this manuscript alongside their explanation are
given by Table~\ref{table:parameters_summary}.

\begin{table}[!htbp]
    \begin{center}
    \resizebox{.7\textwidth}{!}{
    \begin{tabular}{ll}
    \toprule
Feature & Explanation \\
 \midrule
SSE                       & A measure of how far a strategy is from extortionate behaviour defined in~\cite{Knight2019}. \\
$C_{\text{max}}$          & The biggest cooperating rate in the tournament. \\
$C_{\text{min}}$          & The smallest cooperating rate in the tournament. \\
$C_{\text{median}}$       & The median cooperating rate in the tournament. \\
$C_{\text{mean}}$         & The mean cooperating rate in the tournament. \\
$C_r$ / $C_{\text{max}}$    & A strategy's cooperating rate divided by the maximum cooperating rate in the tournament. \\
$C_{\text{min}}$ / $C_r$    & The minimum in the tournament divided by a strategy's cooperating rate. \\
$C_r$ / $C_{\text{median}}$ & A strategy's cooperating rate divided by the median cooperating rate in the tournament. \\
$C_r$ / $C_{\text{mean}}$   & A strategy's cooperating rate divided by the mean cooperating rate in the tournament. \\
$C_r$                       & The cooperating rate of a strategy. \\
$CC$ to $C$ rate            & The probability a strategy will cooperate after a mutual cooperation. \\
$CD$ to $C$ rate            & The probability a strategy will cooperate after being betrayed by the opponent. \\
$DC$ to $C$ rate            & The probability a strategy will cooperate after betraying the opponent. \\
$DD$ to $C$ rate            & The probability a strategy will cooperate after a mutual defection. \\
$p_n$                       & The probability of a player's action being flipped at each interaction. \\
$n$                         & The number of turns in a match. \\
$p_e$                       & The probability of a match ending in the next turn. \\
$N$                         & The number of strategies in the tournament. \\
$k$                         & The number that a given tournament is repeated. \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{The features which are included in the performance evaluation analysis.}\label{table:parameters_summary}
\end{table}