### 2020-03-27 paper/perfomances_table.tex
finalises top performances section

\begin{tabular}{lllllllllllll}
\toprule
{} &                     Name & \$\textbackslash bar\{r\}\$ &                   Name & \$\textbackslash bar\{r\}\$ &                        Name & \$\textbackslash bar\{r\}\$ &               Name & \$\textbackslash bar\{r\}\$ &                       Name & \$\textbackslash bar\{r\}\$ &                 Name & \$\textbackslash bar\{r\}\$ \\
\midrule
0  &            Evolved HMM 5 &     0.007 &                 Grumpy &      0.14 &                         DBS &         0 &          Fortress4 &     0.013 &             Evolved FSM 16 &         0 &           Alternator &     0.304 \\
1  &           Evolved FSM 16 &      0.01 &                    \$e\$ &     0.194 &     Evolved FSM 16 Noise 05 &     0.008 &           Defector &     0.014 &    Evolved FSM 16 Noise 05 &     0.013 &               \$\textbackslash phi\$ &      0.31 \\
2  &     EvolvedLookerUp2\_2\_2 &     0.011 &         Tit For 2 Tats &     0.206 &      Evolved ANN 5 Noise 05 &     0.013 &  Better and Better &     0.016 &                       MEM2 &     0.027 &                  \$e\$ &     0.312 \\
3  &  Evolved FSM 16 Noise 05 &     0.017 &  Slow Tit For Two Tats &      0.21 &                 BackStabber &     0.024 &    Tricky Defector &     0.019 &              Evolved HMM 5 &     0.044 &                \$\textbackslash pi\$ &     0.317 \\
4  &        PSO Gambler 2\_2\_2 &     0.021 &           Cycle Hunter &     0.215 &               DoubleCrosser &     0.025 &          Fortress3 &     0.022 &       EvolvedLookerUp2\_2\_2 &     0.049 &    Limited Retaliate &     0.353 \\
5  &              Evolved ANN &     0.029 &         Risky QLearner &     0.222 &               Evolved ANN 5 &     0.028 &     Gradual Killer &     0.025 &       Spiteful Tit For Tat &      0.06 &     Anti Tit For Tat &     0.354 \\
6  &            Evolved ANN 5 &     0.034 &            Retaliate 3 &     0.229 &                 Evolved ANN &     0.038 &         Aggravater &     0.028 &           Nice Meta Winner &     0.068 &  Limited Retaliate 3 &     0.356 \\
7  &        PSO Gambler 1\_1\_1 &     0.037 &          Cycler CCCCCD &     0.235 &        Spiteful Tit For Tat &     0.051 &             Raider &     0.031 &         NMWE Finite Memory &     0.069 &          Retaliate 3 &     0.356 \\
8  &            Evolved FSM 4 &     0.049 &            Retaliate 2 &     0.239 &               Evolved HMM 5 &     0.051 &         Cycler DDC &     0.045 &         NMWE Deterministic &      0.07 &            Retaliate &     0.357 \\
9  &         PSO Gambler Mem1 &      0.05 &        Defector Hunter &      0.24 &              Level Punisher &     0.052 &        Hard Prober &     0.051 &                    Grudger &      0.07 &          Retaliate 2 &     0.358 \\
10 &                 Winner12 &      0.06 &              Retaliate &     0.242 &                   Omega TFT &     0.059 &         SolutionB1 &      0.06 &           NMWE Long Memory &     0.074 &  Limited Retaliate 2 &     0.361 \\
11 &             Fool Me Once &     0.061 &    Hard Tit For 2 Tats &      0.25 &                Fool Me Once &     0.059 &      Meta Minority &     0.061 &  Nice Meta Winner Ensemble &     0.076 &             Hopeless &     0.368 \\
12 &                      DBS &     0.071 &    Limited Retaliate 3 &     0.253 &  PSO Gambler 2\_2\_2 Noise 05 &     0.067 &              Bully &     0.061 &       EvolvedLookerUp1\_1\_1 &     0.077 &    Arrogant QLearner &     0.407 \\
13 &            DoubleCrosser &     0.072 &               ShortMem &     0.253 &              Evolved FSM 16 &     0.078 &             EasyGo &     0.071 &            NMWE Memory One &      0.08 &    Cautious QLearner &     0.409 \\
14 &              BackStabber &     0.075 &      Limited Retaliate &     0.257 &                  EugineNier &      0.08 &    Fool Me Forever &     0.071 &                   Winner12 &     0.085 &      Fool Me Forever &     0.418 \\
\bottomrule
\end{tabular}