### 2020-03-30 paper/regression_on_median_score.tex
run regression on median score

\section{Multivariate linear regression on median score}\label{app:regression_median_score}

A multivariate linear regression has also been fitted to model the relationship between
the features and the median score. The features included are given
by Table~\ref{table:linear_regression_on_median} alongside their corresponding \(p\) values
in the distinct tournaments and their regression coefficients.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[h]
    \begin{center}
\resizebox{\textwidth}{!}{
\begin{tabular}{lggccggccgg}
    \toprule
    & \multicolumn{2}{g}{Standard} & \multicolumn{2}{c}{Noisy} & \multicolumn{2}{g}{Probabilistic ending} & \multicolumn{2}{c}{Noisy probabilistic ending} & \multicolumn{2}{g}{Overall}\\
    \midrule
    & \multicolumn{2}{g}{\(R\) adjusted: 0.576} & \multicolumn{2}{c}{\(R\) adjusted: 0.679} & \multicolumn{2}{g}{\(R\) adjusted: 0.816} & \multicolumn{2}{c}{\(R\) adjusted: 0.930} & \multicolumn{2}{g}{\(R\) adjusted: 0.318} \\
    {} &  Coefficient &  \(p\)-value &  Coefficient &  \(p\)-value &  Coefficient &  \(p\)-value &  Coefficient &  \(p\)-value &  Coefficient &  \(p\)-value \\
    \midrule
    $CC$ to $C$ rate   &  0.043 &  0.000 & -0.380 &  0.000 &  0.224 &  0.000 &  0.078 &    0.0 &  0.308 &    0.0 \\
    $CD$ to $C$ rate   & -0.325 &  0.000 &  0.124 &  0.000 &  0.060 &  0.000 &  0.073 &    0.0 & -0.014 &    0.0 \\
    $C_r$ / $C_{max}$  &      - &      - & -1.044 &  0.000 &      - &      - & -1.251 &    0.0 &      - &      - \\
    $C_r$ / $C_{mean}$ &  0.553 &  0.000 & -0.101 &  0.000 & -1.136 &  0.000 & -0.089 &    0.0 & -0.665 &    0.0 \\
    $C_{max}$          &  0.059 &  0.000 &      - &      - & -0.044 &  0.086 & -1.396 &    0.0 &      - &      - \\
    $C_{mean}$         &  1.837 &  0.000 &  3.078 &  0.000 &  1.506 &  0.000 &  3.645 &    0.0 &      - &      - \\
    $C_{min}$          &  0.156 &  0.000 &  1.528 &  0.000 &  0.311 &  0.000 &      - &      - &      - &      - \\
    $C_{min}$ / $C_r$  & -0.049 &  0.000 & -0.378 &  0.000 & -0.204 &  0.000 &      - &      - & -0.257 &    0.0 \\
    $DC$ to $C$ rate   & -0.204 &  0.000 &  0.074 &  0.000 &  0.066 &  0.000 &  0.066 &    0.0 &  0.007 &    0.0 \\
    $k$                & -0.000 &  0.853 & -0.000 &  0.987 &  0.000 &  0.008 &  0.000 &    0.0 &      - &      - \\
    $n$                & -0.000 &  0.000 &      - &      - &      - &      - &      - &      - &      - &      - \\
    $p_e$              &      - &      - &      - &      - &  0.025 &  0.000 & -0.095 &    0.0 &      - &      - \\
    $p_n$              &      - &      - &  0.124 &  0.000 &      - &      - &      - &      - &      - &      - \\
    SSE                & -0.294 &  0.000 & -0.319 &  0.000 &  0.055 &  0.000 &  0.010 &    0.0 & -0.015 &    0.0 \\
    constant           &  0.925 &  0.000 &  1.536 &  0.000 &  2.466 &  0.000 &  2.299 &    0.0 &  2.924 &    0.0 \\
    memory usage       &  0.010 &  0.000 & -0.004 &  0.000 &      - &      - &      - &      - &      - &      - \\
    \bottomrule
\end{tabular}}
\end{center}
\caption{Results of multivariate linear regressions with the median score as the dependent variable.
\(R\) squared is reported for each model.}
\label{table:linear_regression_on_median}
\end{table}