### 2020-04-13 paper/correlation_section.tex

\section{Correlation coefficients}\label{app:correlations}

A graphical representation of the correlation coefficients for the features in
Table~\ref{table:manual_features}.

\begin{figure}[!htbp]
        \begin{center}
            \includegraphics[width=.75\linewidth]{../images/standard_correlation_plot.pdf}
        \end{center}
        \caption{Correlation coefficients of features in Table~\ref{table:manual_features}
        for standard tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/noise_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_features}
    for noisy tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/probend_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_features}
    for probabilistic ending tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/probend_noise_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_features}
    for noisy probabilistic ending tournaments}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/merged_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of features in Table~\ref{table:manual_features}
    for data set}
\end{figure}