### 2023-12-01 paper/top_perfomances.tex

\begin{tabular}{lggllggllggllr}
    \toprule
    & \multicolumn{2}{g}{Standard} & \multicolumn{2}{g}{Noisy (\(p_n \leq 0.1\))} & \multicolumn{2}{g}{Probabilistic ending (\(p_e \leq 0.1\))} &  \multicolumn{2}{c}{Noisy probabilistic ending} \\
    \toprule
    {} &                     Name & $\bar{r}$ &                       Name & $\bar{r}$ &                       Name & $\bar{r}$ &                       Name & $\bar{r}$ \\
    \midrule
    0  &            Evolved HMM 5 &     0.007 &                         DBS &       0.0 &             Evolved FSM 16 &       0.0 &                     Raider &     0.022 \\
    1  &           Evolved FSM 16 &      0.01 &     Evolved FSM 16 Noise 05 &     0.008 &    Evolved FSM 16 Noise 05 &     0.013 &                       MEM2 &     0.037 \\
    2  &     EvolvedLookerUp2\_2\_2 &     0.011 &      Evolved ANN 5 Noise 05 &     0.013 &                       MEM2 &     0.027 &                   Prober 3 &     0.039 \\
    3  &  Evolved FSM 16 Noise 05 &     0.017 &                 BackStabber &     0.024 &              Evolved HMM 5 &     0.043 &    Evolved FSM 16 Noise 05 &     0.048 \\
    4  &        PSO Gambler 2\_2\_2 &     0.022 &               DoubleCrosser &     0.025 &       EvolvedLookerUp2\_2\_2 &     0.049 &                Hard Prober &     0.072 \\
    5  &              Evolved ANN &     0.029 &               Evolved ANN 5 &     0.028 &       Spiteful Tit For Tat &     0.059 &       Spiteful Tit For Tat &     0.078 \\
    6  &            Evolved ANN 5 &     0.034 &                 Evolved ANN &     0.038 &           Nice Meta Winner &     0.069 &          Better and Better &     0.089 \\
    7  &        PSO Gambler 1\_1\_1 &     0.037 &        Spiteful Tit For Tat &     0.051 &         NMWE Finite Memory &     0.069 &                    Grudger &     0.091 \\
    8  &            Evolved FSM 4 &     0.049 &               Evolved HMM 5 &     0.051 &         NMWE Deterministic &      0.07 &                  Fortress4 &     0.096 \\
    9  &         PSO Gambler Mem1 &      0.05 &              Level Punisher &     0.052 &                    Grudger &      0.07 &     Meta Winner Memory One &     0.099 \\
    10 &                 Winner12 &      0.06 &                   Omega TFT &     0.059 &           NMWE Long Memory &     0.074 &           NMWE Long Memory &     0.099 \\
    11 &             Fool Me Once &     0.061 &                Fool Me Once &     0.059 &  Nice Meta Winner Ensemble &     0.076 &           Nice Meta Winner &     0.104 \\
    12 &                      DBS &     0.071 &  PSO Gambler 2\_2\_2 Noise 05 &     0.067 &       EvolvedLookerUp1\_1\_1 &     0.077 &         NMWE Deterministic &     0.109 \\
    13 &            DoubleCrosser &     0.072 &              Evolved FSM 16 &     0.078 &            NMWE Memory One &      0.08 &            NMWE Memory One &     0.112 \\
    14 &              BackStabber &     0.075 &                  EugineNier &      0.08 &            NMWE Stochastic &     0.085 &  Nice Meta Winner Ensemble &     0.115 \\
    \bottomrule
    \end{tabular}