### 2023-12-01 paper/correlation_section.tex
finish SI section analysis with all data and clean notebooks
The notebooks now include code to run the analysis for both the
main paper and the SI

\section{Correlation coefficients}\label{app:correlations}

In this section we present the correlation coefficients for the features
presented in Table 3 of the main manuscript. The correlation coefficients
are calculated using the Spearman's rank correlation coefficient. The results
are presented as a graphical representation.

\begin{figure}[!htbp]
        \begin{center}
            \includegraphics[width=.75\linewidth]{../images/standard_correlation_plot.pdf}
        \end{center}
        \caption{Correlation coefficients of strategies' features for standard tournaments.}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/subset_noise_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of strategies' features for noisy tournaments with $p_n \leq 0.1$.}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/noise_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of strategies' features for noisy tournaments.}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/subset_probend_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of strategies' features for probabilistic ending tournaments with $p_e \leq 0.1$.}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/probend_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of strategies' features for probabilistic ending tournaments.}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/subset_probend_noise_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of strategies' features for noisy probabilistic ending tournaments
    with $p_n \leq 0.1$ and $p_e \leq 0.1$.}
\end{figure}
\begin{figure}[!htbp]
    \begin{center}
        \includegraphics[width=.75\linewidth]{../images/probend_noise_correlation_plot.pdf}
    \end{center}
    \caption{Correlation coefficients of strategies' features for noisy probabilistic ending tournaments.}
\end{figure}