### 2023-12-01 paper/si.tex

\documentclass{article}
\usepackage[margin=2.5cm, includefoot, footskip=30pt]{geometry}

\setlength{\parindent}{0em}
\setlength{\parskip}{1em}
\renewcommand{\baselinestretch}{1}

%%%%Packages%%%%
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{graphics}
\usepackage{multicol}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{color,colortbl}
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{wrapfig, blindtext}
\usepackage{soul}
\usepackage[table]{xcolor}
%%%%%%%%%%%%%%%%%

\definecolor{Gray}{gray}{0.92}
\usepackage[first=0,last=9]{lcg}
\newcommand{\ra}{\rand0.\arabic{rand}}

\newcommand{\uniquenumberofseeds}{\input{unique_number_of_seeds.tex}}
\newcommand{\numberofalltournaments}{\input{number_of_all_tournaments.tex}}
\newcommand{\numberofstrategies}{\input{unique_number_of_strategies.tex}}
\def\axelrod{\texttt{Axelrod-Python }}

\setlength{\tabcolsep}{3pt}

\title{{\LARGE Electronic supplementary material}\\{\bf \LARGE Properties of Winning Iterated Prisoner's Dilemma Strategies}}
\author{Nikoleta E. Glynatsi, Vincent A. Knight, Marc Harper}
\date{}

\begin{document}
\maketitle

%% BRIEF INTRO %%

\noindent
This document provides additional details and analysis based on the complete
datasets for noisy and probabilistic ending tournaments. In
Section~\ref{app:parameters}, a summary of all parameters and notation used in
the manuscript is presented. Section~\ref{app:analysis_full_dataset} offers
further details on the correlation analysis and includes additional information.
In Section~\ref{app:correlations}, we present the results of the correlation
analysis when all features of the strategies are considered. In
Section~\ref{app:clustering}, the results of the random forest and clustering
evaluations with the inclusion of all features are discussed.
Section~\ref{app:regression_median_score} presents the results of the regression
analysis on the median score of the strategies. Finally, in
Section~\ref{app:list_of_players}, a comprehensive list of all strategies
considered in this work is provided.

\input{parameters_section}
\input{result_with_entire_dataset}
\input{correlation_section}
\input{random_forest_evaluation}
\input{regression_on_median_score}
\input{strategies_list_section}

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}