### 2025-07-28 paper/top_perfomances_full_data_set.tex

\begin{tabular}{lggllggllggllr}
    \toprule
    & \multicolumn{2}{g}{Standard} & \multicolumn{2}{c}{Noisy} &  \multicolumn{2}{g}{Noisy (\(p_n < 0.1\))} & \multicolumn{2}{c}{Probabilistic ending} & \multicolumn{2}{g}{Probabilistic ending (\(p_e < 0.1\))} &  \multicolumn{2}{c}{Noisy probabilistic ending} \\
    \midrule
    & Name & $\bar{r}$ &                 Name & $\bar{r}$ &               Name & $\bar{r}$ &                 Name & $\bar{r}$ &                 Name & $\bar{r}$ &                 Name & $\bar{r}$ \\
    \midrule
    0  &            Evolved HMM 5 &     0.007 &                 Grumpy &      0.14 &                         DBS &       0.0 &          Fortress4 &     0.013 &             Evolved FSM 16 &       0.0 &           Alternator &     0.304 \\
    1  &           Evolved FSM 16 &      0.01 &                    \$e\$ &      0.19 &     Evolved FSM 16 Noise 05 &     0.008 &           Defector &     0.014 &    Evolved FSM 16 Noise 05 &     0.013 &               \$\textbackslash phi\$ &      0.31 \\
    2  &     EvolvedLookerUp2\_2\_2 &     0.011 &         Tit For 2 Tats &     0.206 &      Evolved ANN 5 Noise 05 &     0.013 &  Better and Better &     0.016 &                       MEM2 &     0.027 &                  \$e\$ &     0.312 \\
    3  &  Evolved FSM 16 Noise 05 &     0.017 &  Slow Tit For Two Tats &      0.21 &                 BackStabber &     0.024 &    Tricky Defector &     0.019 &              Evolved HMM 5 &     0.043 &                \$\textbackslash pi\$ &     0.317 \\
    4  &        PSO Gambler 2\_2\_2 &     0.022 &           Cycle Hunter &     0.215 &               DoubleCrosser &     0.025 &          Fortress3 &     0.022 &       EvolvedLookerUp2\_2\_2 &     0.049 &    Limited Retaliate &     0.353 \\
    5  &              Evolved ANN &     0.029 &         Risky QLearner &     0.222 &               Evolved ANN 5 &     0.028 &     Gradual Killer &     0.025 &       Spiteful Tit For Tat &     0.059 &     Anti Tit For Tat &     0.354 \\
    6  &            Evolved ANN 5 &     0.034 &          Cycler CCCCCD &     0.229 &                 Evolved ANN &     0.038 &         Aggravater &     0.028 &           Nice Meta Winner &     0.069 &  Limited Retaliate 3 &     0.356 \\
    7  &        PSO Gambler 1\_1\_1 &     0.037 &            Retaliate 3 &      0.23 &        Spiteful Tit For Tat &     0.051 &             Raider &     0.031 &         NMWE Finite Memory &     0.069 &          Retaliate 3 &     0.356 \\
    8  &            Evolved FSM 4 &     0.049 &            Retaliate 2 &     0.239 &               Evolved HMM 5 &     0.051 &         Cycler DDC &     0.045 &         NMWE Deterministic &      0.07 &            Retaliate &     0.357 \\
    9  &         PSO Gambler Mem1 &      0.05 &        Defector Hunter &      0.24 &              Level Punisher &     0.052 &        Hard Prober &     0.051 &                    Grudger &      0.07 &          Retaliate 2 &     0.358 \\
    10 &                 Winner12 &      0.06 &              Retaliate &     0.242 &                   Omega TFT &     0.059 &         SolutionB1 &      0.06 &           NMWE Long Memory &     0.074 &  Limited Retaliate 2 &     0.361 \\
    11 &             Fool Me Once &     0.061 &    Hard Tit For 2 Tats &      0.25 &                Fool Me Once &     0.059 &      Meta Minority &     0.061 &  Nice Meta Winner Ensemble &     0.076 &             Hopeless &     0.368 \\
    12 &                      DBS &     0.071 &      Arrogant QLearner &      0.25 &  PSO Gambler 2\_2\_2 Noise 05 &     0.067 &              Bully &     0.061 &       EvolvedLookerUp1\_1\_1 &     0.077 &    Arrogant QLearner &     0.406 \\
    13 &            DoubleCrosser &     0.072 &    Limited Retaliate 3 &     0.253 &              Evolved FSM 16 &     0.078 &             EasyGo &     0.071 &            NMWE Memory One &      0.08 &    Cautious QLearner &     0.409 \\
    14 &              BackStabber &     0.075 &               ShortMem &     0.253 &                  EugineNier &      0.08 &    Fool Me Forever &     0.071 &            NMWE Stochastic &     0.085 &      Fool Me Forever &     0.418 \\
    \bottomrule
    \end{tabular}